Measuring matter effects in Neutron star - Black hole binaries with Advanced LIGO

### 2015-11-13 Paper/paper.tex
added figures showing TN/NN results

\begin{figure*}
\centering 
\includegraphics[trim={{0.2\columnwidth} 0 0 0},width=2.1\columnwidth]{plots-TN/TN_MchirpBiases_Lambda_SNR.pdf}\\ 
\caption{These figures show the (fractional) difference between the chirp mass 
value corresponding to the median of its posterior probability distribution,
and the actual injected chirp mass,
as a function of the signal-to-noise-ratio (SNR) of the injected signal.
% In the top row, NS matter effects on the inspiral and merger waveform
% are neithed included in templates nor in the injected signal. 
In the top row, NS matter effects (inspiral and merger) are included in the
injections but not in templates. This corresponds to the current 
plan for aLIGO parameter estimation studies. The bottom row shows the effect
of additionally including NS matter effects in the templates.

@@

In each row, the left and right panel correspond to 
$q=m_\mathrm{BH}/m_\mathrm{NS}=\{2,3\}$, respectively. 
In each panel, the color of each curve corresponds to the value of NS 
deformability parameter $(\Lambda=\{500,1000,2000\})$, 
and line-style corresponds to the value of
the dimensionless spin on the BH ($\chi_\mathrm{BH}=\{-0.5,0,+0.5\}$).
}
\label{fig:TNT_chirpMassBias_vs_SNR_q23}
\end{figure*}
\begin{figure*}
\centering
\includegraphics[trim={{0.4\columnwidth} 0 0 0},width=2.2\columnwidth]{plots-TN/TN_MchirpCIWidths90_0_Lambda_SNR.pdf}
\caption{These figures are similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23}
with the difference that here we show the width of the $90\%$ confidence 
interval (recovered) for chirp mass, as a function of the injected SNR. 
Note that the confidence interval's width is normalized by the injected
parameter value.
}
\label{fig:TNT_chirpMassCIWidth90_vs_SNR_q23}
\end{figure*}
\newpage
\newpage

\begin{figure*}[!t]
\centering    
\textbf{NS-BH Injection; BH-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TN_EtaBias_vs_SNR_q23.pdf}\\ 
\textbf{NS-BH Injection; NS-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TT_EtaBias_vs_SNR_q23.pdf}\\%\quad
\textbf{BH-BH Injection; BH-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/NN_EtaBias_vs_SNR_q23.pdf} 
\caption{This figure (top two rows) is similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23},
with the difference that we consider the systematic bias in the dimensionless
mass-ratio $\eta$ here. In the bottom row we show control runs where both
signal and template waveforms lack matter effects.}
\label{fig:TNT_EtaBias_vs_SNR_q23}
\end{figure*}
% 
\begin{figure*}
\centering    
\textbf{NS-BH Injection; BH-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TN_EtaCIWidth90_vs_SNR_q23.pdf}\\ 
\textbf{NS-BH Injection; NS-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TT_EtaCIWidth90_vs_SNR_q23.pdf}\\%\quad
\textbf{BH-BH Injection; BH-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/NN_EtaCIWidth90_vs_SNR_q23.pdf} 
\caption{This figure (top two rows) is similar to Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23},
  with the difference that we consider the confidence intervals in the recovery
  of the dimensionless mass-ratio $\eta$ here. The bottom row here is similar to 
  Fig.~\ref{fig:TNT_EtaBias_vs_SNR_q23} with the difference that width of the $90\%$
  confidence interval is shown here.}
\label{fig:TNT_EtaCIWidth90_vs_SNR_q23}
\end{figure*}
% 
\begin{figure*}
\centering    
\textbf{NS-BH Injection; BH-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TN_BHspinBias_vs_SNR_q23.pdf}\\ 
\textbf{NS-BH Injection; NS-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TT_BHspinBias_vs_SNR_q23.pdf}\\%\quad
\textbf{BH-BH Injection; BH-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/NN_BHspinBias_vs_SNR_q23.pdf}
\caption{The top two rows in this figure are
similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23},
with the differences that we consider the systematic bias in the dimensionless
spin on the BH here, and that the differences shown are absolute and not
fractions of the injected spin. In the bottom row, we show a control run where
both injected and template waveforms correspond to binary black holes.}
\label{fig:TNT_BHspinBias_vs_SNR_q23}
\end{figure*}
% 
\begin{figure*}
\centering    
\textbf{BH-BH Injection; BH-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/NN_BHspinCIWidth90_vs_SNR_q23.pdf}\\ 
\textbf{NS-BH Injection; BH-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TN_BHspinCIWidth90_vs_SNR_q23.pdf}\\ 
\textbf{NS-BH Injection; NS-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TT_BHspinCIWidth90_vs_SNR_q23.pdf}
\caption{The top two rows in the figure are similar to
Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23},
with the difference that we consider the confidence intervals in the recovery
of the dimensionless spin on the BH $\chi_\mathrm{BH}$ here.
As in Fig.~\ref{fig:TNT_BHspinBias_vs_SNR_q23}, we show here the absolute 
width of the confidence interval without normalizing with the injected spin.
The bottom row show a control run where matter effects are ignored both 
in the injected signal and the template waveforms.}
\label{fig:TNT_BHspinCIWidth90_vs_SNR_q23}
\end{figure*}
% 
% 
Advanced LIGO searches and parameter estimation efforts aimed at binaries 
containing NS and stellar-mass BHs are poised to use waveform models that do not
include the effects of the NS tidal deformability~\cite{Canton:2014ena}.
While this is not expected to
be the dominant source of error at low SNRs, it will likely introduce a
systematic bias in the physical parameters that we recover from GW observations.
In this section, we study the impact of the same for binaries where the BH spins
are aligned to the orbital angular momentum.

>>>>

@@

% \begin{figure*}[!t]
% \includegraphics[width=1.7\columnwidth]{plots/TN_EtaBias_vs_SNR_q23.pdf}\\ 
% \includegraphics[width=1.7\columnwidth]{plots/TT_EtaBias_vs_SNR_q23.pdf}\\%\quad
% \textbf{BH-BH Injection; BH-BH Templates}
% \includegraphics[width=1.7\columnwidth]{plots/NN_EtaBias_vs_SNR_q23.pdf} 
% \caption{This figure (top two rows) is similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23},
% with the difference that we consider the systematic bias in the dimensionless
% mass-ratio $\eta$ here. In the bottom row we show control runs where both
% signal and template waveforms lack matter effects.}
% \label{fig:TNT_EtaBias_vs_SNR_q23}
% \includegraphics[width=1.7\columnwidth]{plots/TN_EtaCIWidth90_vs_SNR_q23.pdf}\\ 
% \includegraphics[width=1.7\columnwidth]{plots/TT_EtaCIWidth90_vs_SNR_q23.pdf}\\%\quad
% \textbf{BH-BH Injection; BH-BH Templates}
% \includegraphics[width=1.7\columnwidth]{plots/NN_EtaCIWidth90_vs_SNR_q23.pdf} 
% \caption{This figure (top two rows) is similar to Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23},
%   with the difference that we consider the confidence intervals in the recovery
%   of the dimensionless mass-ratio $\eta$ here. The bottom row here is similar to 
%   Fig.~\ref{fig:TNT_EtaBias_vs_SNR_q23} with the difference that width of the $90\%$
%   confidence interval is shown here.}
% \label{fig:TNT_EtaCIWidth90_vs_SNR_q23}
% \includegraphics[width=1.7\columnwidth]{plots/TN_BHspinBias_vs_SNR_q23.pdf}\\ 
% \includegraphics[width=1.7\columnwidth]{plots/TT_BHspinBias_vs_SNR_q23.pdf}\\%\quad
% \textbf{BH-BH Injection; BH-BH Templates}
% \includegraphics[width=1.7\columnwidth]{plots/NN_BHspinBias_vs_SNR_q23.pdf}
% \caption{The top two rows in this figure are
% similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23},
% with the differences that we consider the systematic bias in the dimensionless
% spin on the BH here, and that the differences shown are absolute and not
% fractions of the injected spin. In the bottom row, we show a control run where
% both injected and template waveforms correspond to binary black holes.}
% \label{fig:TNT_BHspinBias_vs_SNR_q23}
% \textbf{BH-BH Injection; BH-BH Templates}
% \includegraphics[width=1.7\columnwidth]{plots/NN_BHspinCIWidth90_vs_SNR_q23.pdf}\\ 
% \includegraphics[width=1.7\columnwidth]{plots/TN_BHspinCIWidth90_vs_SNR_q23.pdf}\\ 
% \includegraphics[width=1.7\columnwidth]{plots/TT_BHspinCIWidth90_vs_SNR_q23.pdf}
% \caption{The top two rows in the figure are similar to
% Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23},
% with the difference that we consider the confidence intervals in the recovery
% of the dimensionless spin on the BH $\chi_\mathrm{BH}$ here.
% As in Fig.~\ref{fig:TNT_BHspinBias_vs_SNR_q23}, we show here the absolute 
% width of the confidence interval without normalizing with the injected spin.
% The bottom row show a control run where matter effects are ignored both 
% in the injected signal and the template waveforms.}
% \label{fig:TNT_BHspinCIWidth90_vs_SNR_q23}

% #################
\begin{figure*}
\centering 
\includegraphics[trim={{0.2\columnwidth} 0 0 0},width=2.1\columnwidth]{plots-TN/TN_MchirpBiases_Lambda_SNR.pdf}
\caption{These figures show the (fractional) difference between the chirp mass 
value corresponding to the median of its posterior probability distribution,
and the actual injected chirp mass,
as a function of the signal-to-noise-ratio (SNR) of the injected signal.
% In the top row, NS matter effects on the inspiral and merger waveform
% are neithed included in templates nor in the injected signal. 
In the top row, NS matter effects (inspiral and merger) are included in the
injections but not in templates. This corresponds to the current 
plan for aLIGO parameter estimation studies. The bottom row shows the effect
of additionally including NS matter effects in the templates.
% 
In each row, the left and right panel correspond to 
$q=m_\mathrm{BH}/m_\mathrm{NS}=\{2,3\}$, respectively. 
In each panel, the color of each curve corresponds to the value of NS 
deformability parameter $(\Lambda=\{500,1000,2000\})$, 
and line-style corresponds to the value of
the dimensionless spin on the BH ($\chi_\mathrm{BH}=\{-0.5,0,+0.5\}$).
}
\label{fig:TNT_chirpMassBias_vs_SNR_q23}
\end{figure*}
% 
\begin{figure*}
\centering
\includegraphics[trim={{0.4\columnwidth} 0 0 0},width=2.2\columnwidth]{plots-TN/TN_MchirpCIWidths90_0_Lambda_SNR.pdf}
\caption{These figures are similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23}
with the difference that here we show the width of the $90\%$ confidence 
interval (recovered) for chirp mass, as a function of the injected SNR. 
Note that the confidence interval's width is normalized by the injected
parameter value.
}
\label{fig:TNT_chirpMassCIWidth90_vs_SNR_q23}
\end{figure*}
% 
\newpage
\newpage

\begin{figure*}[!t]
\centering    
\includegraphics[trim={{0.4\columnwidth} 0 0 0},width=2.2\columnwidth]{plots-TN/TN_EtaBiases_Lambda_SNR.pdf}
\caption{This figure (top two rows) is similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23},
with the difference that we consider the systematic bias in the dimensionless
mass-ratio $\eta$ here. In the bottom row we show control runs where both
signal and template waveforms lack matter effects.}
\label{fig:TNT_EtaBias_vs_SNR_q23}
\end{figure*}
% 
\begin{figure*}
\centering    
\includegraphics[trim={{0.4\columnwidth} 0 0 0},width=2.2\columnwidth]{plots-TN/TN_EtaCIWidths68_3_Lambda_SNR.pdf}
\caption{This figure is similar to
Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23},
  with the difference that we consider the confidence intervals in the recovery
  of the dimensionless mass-ratio $\eta$ here. The bottom row here is similar to 
  Fig.~\ref{fig:TNT_EtaBias_vs_SNR_q23} with the difference that width of the $90\%$
  confidence interval is shown here.}
\label{fig:TNT_EtaCIWidth90_vs_SNR_q23}
\end{figure*}
% 
\begin{figure*}
\centering
\includegraphics[trim={{0.4\columnwidth} 0 0 0},width=2.2\columnwidth]{plots-TN/TN_ChiBHBiases_CI90_0_Lambda_SNR.pdf}
\caption{The top two rows in this figure are
similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23},
with the differences that we consider the systematic bias in the dimensionless
spin on the BH here, and that the differences shown are absolute and not
fractions of the injected spin. In the bottom row, we show a control run where
both injected and template waveforms correspond to binary black holes.}
\label{fig:TNT_BHspinBias_vs_SNR_q23}
\end{figure*}
% 
\begin{figure*}
\centering    
\includegraphics[trim={{0.4\columnwidth} 0 0 0},width=2.2\columnwidth]{plots-TN/TN_ChiBHCIWidths90_0_Lambda_SNR.pdf}
\caption{The top two rows in the figure are similar to
Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23},
with the difference that we consider the confidence intervals in the recovery
of the dimensionless spin on the BH $\chi_\mathrm{BH}$ here.
As in Fig.~\ref{fig:TNT_BHspinBias_vs_SNR_q23}, we show here the absolute 
width of the confidence interval without normalizing with the injected spin.
The bottom row show a control run where matter effects are ignored both 
in the injected signal and the template waveforms.}
\label{fig:TNT_BHspinCIWidth90_vs_SNR_q23}
\end{figure*}
% 
% 
Advanced LIGO searches and parameter estimation efforts aimed at binaries 
containing NS and stellar-mass BHs are poised to use waveform models that do not
include the effects of the NS tidal deformability~\cite{Canton:2014ena}.
While this is not expected to
be the dominant source of error at low SNRs, it will likely introduce a
systematic bias in the physical parameters that we recover from GW observations.
In this section, we study the impact of the same for binaries where the BH spins
are aligned to the orbital angular momentum.
